---
title: "Stat 184 Final Project"
output: html_document
date: "2024-12-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)

data <- read.csv("cleaned_scout_data.csv")

model_counts <- data %>% 
  group_by(make_model) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))

top_3_models <- head(model_counts, 3)

library(ggplot2)
ggplot(data, aes(x = make_model, fill = make_model)) +
  geom_bar(color = "black") +
  geom_text(data = top_3_models, aes(x = make_model, y = count, label = count), 
            vjust = -0.5, color = "black", size = 3) +
  labs(
    title = "Histogram of All Car Models",
    x = "Car Model",
    y = "Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


***Summary table for car types we interest in***
```{r}
#Provides key statistics for price, mileage, and age
summary_table <- data %>%
  filter(make_model %in% c("Audi A3", "Opel Insignia", "Audi A1")) %>%
  group_by(make_model) %>%
  summarise(
    Avg_Price = mean(price, na.rm = TRUE),
    Median_Price = median(price, na.rm = TRUE),
    Min_Price = min(price, na.rm = TRUE),
    Max_Price = max(price, na.rm = TRUE),
    Avg_Mileage = mean(mileage_km, na.rm = TRUE),
    Median_Mileage = median(mileage_km, na.rm = TRUE),
    Min_Mileage = min(mileage_km, na.rm = TRUE),
    Max_Mileage = max(mileage_km, na.rm = TRUE),
    Avg_Age = mean(age, na.rm = TRUE),
    Median_Age = median(age, na.rm = TRUE),
    Min_Age = min(age, na.rm = TRUE),
    Max_Age = max(age, na.rm = TRUE)
  )

print(summary_table)
```



***Price vs Age***

**Audi A3 Price vs Age**
```{r}
audi_a3_data <- data %>% filter(make_model == "Audi A3")

# Scatter plot with regression line
ggplot(audi_a3_data, aes(x = age, y = price)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  labs(
    title = "Price vs Age for Audi A3",
    x = "Age (years)",
    y = "Price (in units)"
  ) +
  theme_minimal()
audi_age_model <- lm(price ~ age, data = audi_a3_data)
summary(audi_age_model)


# Boxplot
ggplot(audi_a3_data, aes(x = factor(age), y = price)) +
  geom_boxplot(fill = "pink") +
  labs(
    title = "Boxplot of Price by Age for Audi A3",
    x = "Age (years)",
    y = "Price (in units)"
  ) +
  theme_minimal()
```


**Opel Insignia Price vs Age**

```{r}
opel_insignia_data <- data %>% filter(make_model == "Opel Insignia")

# Scatter plot with regression line
ggplot(opel_insignia_data, aes(x = age, y = price)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(
    title = "Price vs Age for Opel Insignia",
    x = "Age (years)",
    y = "Price (in units)"
  ) +
  theme_minimal()

opel_age_model <- lm(price ~ age, data = opel_insignia_data)
summary(opel_age_model)

# Boxplot
ggplot(opel_insignia_data, aes(x = factor(age), y = price)) +
  geom_boxplot(fill = "lightblue") +
  labs(
    title = "Boxplot of Price by Age for Opel Insignia",
    x = "Age (years)",
    y = "Price (in units)"
  ) +
  theme_minimal()

```
#scatter plot：This plot shows the relationship between age and price for Opel Insignia cars
#Boxplot: This plot shows the distribution of price for each age group


**Audi A1 Price vs Age**
```{r}
audi_a1_data <- data %>% filter(make_model == "Audi A1")

# Scatter plot with regression line
ggplot(audi_a1_data, aes(x = age, y = price)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "green") +
  labs(
    title = "Price vs Age for Audi A1",
    x = "Age (years)",
    y = "Price (in units)"
  ) +
  theme_minimal()

audi1_age_model <- lm(price ~ age, data = audi_a1_data)
summary(audi1_age_model)

# Boxplot for price by age for Audi A1
ggplot(audi_a1_data, aes(x = factor(age), y = price)) +
  geom_boxplot(fill = "lightgreen") +
  labs(
    title = "Boxplot of Price by Age for Audi A1",
    x = "Age (years)",
    y = "Price (in units)"
  ) +
  theme_minimal()
```

#Scatter plot: This plot shows the relationship between age and price for Audi A1 cars.
#Boxplot: This plot shows the distribution of price for each age group


***Price vs Mileage***

**Audi A3 Price vs Mileage**

```{r}
#scatter plot
ggplot(audi_a3_data, aes(x = mileage_km, y = price)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  scale_x_continuous(labels = scales::comma, breaks = seq(0, 300000, by = 50000)) +
  labs(
    title = "Price vs Mileage for Audi A3",
    x = "Mileage (km)",
    y = "Price (in units)"
  ) +
  theme_minimal()

audi_mile_model <- lm(price ~ mileage_km, data = audi_a3_data)
summary(audi_mile_model)

# Boxplot
ggplot(audi_a3_data, aes(x = cut(mileage_km, breaks = c(0, 50000, 100000, 150000, 200000, 250000, 300000),
                                labels = c("0-50k", "50k-100k", "100k-150k", "150k-200k", "200k-250k", "250k-300k")),
                        y = price)) +
  geom_boxplot(fill = "pink") +
  labs(
    title = "Boxplot of Price by Mileage for Audi A3",
    x = "Mileage Range (km)",
    y = "Price (in units)"
  ) +
  theme_minimal()
```

**Opel Insignia Price vs Mileage**

```{r}
#Scatter plot
ggplot(opel_insignia_data, aes(x = mileage_km, y = price)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  scale_x_continuous(labels = scales::comma, breaks = seq(0, 300000, by = 50000)) +
  labs(
    title = "Price vs Mileage for Opel Insignia",
    x = "Mileage (km)",
    y = "Price (in units)"
  ) +
  theme_minimal()

opel_age_model <- lm(price ~ mileage_km, data = opel_insignia_data)
summary(opel_age_model)

# Boxplot
ggplot(opel_insignia_data, aes(x = cut(mileage_km, breaks = c(0, 50000, 100000, 150000, 200000, 250000, 300000),
                                     labels = c("0-50k", "50k-100k", "100k-150k", "150k-200k", "200k-250k", "250k-300k")),
                             y = price)) +
  geom_boxplot(fill = "lightblue") +
  labs(
    title = "Boxplot of Price by Mileage for Opel Insignia",
    x = "Mileage Range (km)",
    y = "Price (in units)"
  ) +
  theme_minimal()
```

**Audi A1**

```{r}
#Scatter plot
ggplot(audi_a1_data, aes(x = mileage_km, y = price)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "green") +
  scale_x_continuous(labels = scales::comma, breaks = seq(0, 300000, by = 50000)) +
  labs(
    title = "Price vs Mileage for Audi A1",
    x = "Mileage (km)",
    y = "Price (in units)"
  ) +
  theme_minimal()

audi1_age_model <- lm(price ~ mileage_km, data = audi_a1_data)
summary(audi1_age_model)

# Boxplot
# This plot shows the distribution of price for each mileage group
ggplot(audi_a1_data, aes(x = cut(mileage_km, breaks = c(0, 50000, 100000, 150000, 200000, 250000, 300000),
                                labels = c("0-50k", "50k-100k", "100k-150k", "150k-200k", "200k-250k", "250k-300k")),
                        y = price)) +
  geom_boxplot(fill = "lightgreen") +
  labs(
    title = "Boxplot of Price by Mileage for Audi A1",
    x = "Mileage Range (km)",
    y = "Price (in units)"
  ) +
  theme_minimal()


```






